Aquifer depletion exacerbates agricultural drought losses in the US High Plains

### 2023-09-02 Writing/NatureWater/2ndRoundRevision/replies/reply_to_reviewers_R2.tex

\textcolor{blue}{Thank you for catching this. Fixed.}

\item Line 428 – is ``a linear in parameter model'' correct? Wouldn't ``a linear parameter model'' be better?

\textcolor{blue}{We rewrote it to the following:}

\textcolor{blue}{``Notice that unlike non-parametric regression, GAM is actually a linear (in-parameter) model and estimate $\beta_k$ to fit the data best.''}

@@

\textcolor{blue}{In statistics, ``linear'' model means that the model is linear in parameter, but not necessarily in independent variable. For example, the model below is linear in parameter, but not linear in $x$.}

\begin{align}
  y = \beta x^2 + v
\end{align}

\textcolor{blue}{ GAM is linear in parameter, but not linear in independent variable. We wanted to emphasize the distinction instead of simply saying linear model.}

@@

\textcolor{blue}{We changed all the occurrences of $\alpha_{k}$ to $\eta_{k}$.}

\item (M2.c) Lines 427-428: Please clarify the grammatical structure of the following sentence: ``Notice that unlike non-parametric regression, GAM is actually a linear in parameter model and estimate $\beta_k$ to fit the data best.''

\textcolor{blue}{We rewrote it to the following:}

\textcolor{blue}{``Notice that unlike non-parametric regression, GAM is actually a linear (in parameter) model and estimate $\beta_k$ to fit the data best.''}

### 2023-09-02 Writing/NatureWater/2ndRoundRevision/manuscript.tex
Edits to cover letter.
minor text change

where county and year are indicated by subscripts \(i\) and \(t\), respectively, \(y\) is crop yield, and \(WB\) is seasonal water deficit. \(C_j\) is a dummy variable that takes 1 if observation \(i\) is under category \(C_j\), 0 otherwise. The observations are grouped into four categories: rainfed production, and county-years with aquifer thickness falling in the \([0\%-33.3\%)\), \([33.3\%-66.7\%)\), and \([66.7\% - 100\%]\) quantile ranges of aquifer thickness. The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category. Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) representing the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (see Appendix \ref{spline-basis} for further explanations on how spline-based regression works). $K$ is the total number of spline basis functions. $K = 3$ was chosen for this regression as it minimizes BIC. To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\lambda_t\)) are also included and standard error estimation is clustered by county.

@@

Different coefficients (or namely weights) given to the spline basis functions result in different shapes of the impact of \(x\) on \(y\). Figure \ref{fig:ex-fx} presents two examples of two different sets of coefficients and their resulting representation of \(f(x)\). In the first case, \(\{\beta_1, \dots, \beta_5\} = \{60, 60, 60, 60, 60\}\). In the second case, \(\{\beta_1, \dots, \beta_5\} = \{20, -20, -40, 40, 60\}\). Among all the possibilities of the combinations of the coefficients, GAM finds the set of coefficient values that fits the data best. Notice that unlike non-parametric regression, GAM is actually a linear-in-parameter model and estimate \(\beta_k\) to fit the data best.

>>>>

where county and year are indicated by subscripts \(i\) and \(t\), respectively, \(y\) is crop yield, and \(WB\) is seasonal water deficit. \(C_j\) is a dummy variable that takes 1 if observation \(i\) is under category \(C_j\), 0 otherwise. The observations are grouped into four categories: rainfed production, and county-years with aquifer thickness falling in the \([0\%-33.3\%)\), \([33.3\%-66.7\%)\), and \([66.7\% - 100\%]\) quantile ranges of aquifer thickness. The three irrigated production groups have about the same number of observations, which ensures statistical stability in estimating yield response to drought by category. Spline functions are used to capture the potential non-linearity of the impact of water deficit on yield in a flexible manner, with \(\sigma_j\) representing the category specific intercept, \(\phi_{j,k}\) \(k\)th spline basis function, and \(\beta_k\) is its coefficient (see Appendix \ref{spline-basis} for further explanations on how spline-based regression works). $K$ is the total number of spline basis functions. $K = 3$ was chosen for this regression as it minimizes Bayesian Information Criterion (BIC). To control for time-invariant heterogeneity across counties and yearly shocks, county fixed effects (\(\alpha_i\)) and year fixed effects (\(\lambda_t\)) are also included and standard error estimation is clustered by county.

@@

Different coefficients (or namely weights) given to the spline basis functions result in different shapes of the impact of \(x\) on \(y\). Figure \ref{fig:ex-fx} presents two examples of two different sets of coefficients and their resulting representation of \(f(x)\). In the first case, \(\{\beta_1, \dots, \beta_5\} = \{60, 60, 60, 60, 60\}\). In the second case, \(\{\beta_1, \dots, \beta_5\} = \{20, -20, -40, 40, 60\}\). Among all the possibilities of the combinations of the coefficients, GAM finds the set of coefficient values that fits the data best. Notice that unlike non-parametric regression, GAM is actually a linear (in parameter) model and estimate \(\beta_k\) to fit the data best.
